---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
library(tidyverse)
deaths <- av %>% pivot_longer(cols = starts_with("Death"), names_to = "Time", values_to = "Died")

deaths <- deaths %>% mutate(
  Time = parse_number(Time))

deaths = deaths %>% filter(Died != "") %>% group_by(URL, Died) %>% mutate(Time = max(Time, na.rm = TRUE))

deaths <- deaths %>% unique()

meanDeath <- mean(deaths$Time)

```

**We observe that there are 1.28 deaths on average. **
 
Similarly, deal with the returns of characters.

```{r}
returns <- av %>% pivot_longer(cols = starts_with("Return"), names_to = "Time", values_to = "Returned")

returns <- returns %>% mutate(
  Time = parse_number(Time))

returns = returns %>% filter(Returned != "") %>% group_by(URL, Returned) %>% mutate(Time = max(Time, na.rm = TRUE))

returns <- returns %>% unique()

meanReturn <- mean(returns$Time)

```

**We observe that there are 1.49 returns on average. **




## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

Given the Avengers’ 53 years in operation and overall mortality rate, fans of the comics can expect one current or former member to die every seven months or so, with a permanent death occurring once every 20 months.

### Include the code

```{r}
total_deaths <- deaths %>%
  filter(Death == "yes") %>%
  nrow()

years <- 53

average_death_frequency <- total_deaths / years


permanent_deaths <- deaths %>%
  filter(Death == "yes" & Return == "no") %>%
  nrow()

average_permanent_death_frequency <- permanent_deaths / years


average_death_frequency
average_permanent_death_frequency


```

FiveThirtyEight statement:
"Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team"

```{r}
#Vinayak Trigunayat
numAvengers <- nrow(av)
avengersWhoDied <- deaths %>% filter(Died == "YES")
numAvengersDied <- nrow(avengersWhoDied)
```
**We see that this result is true, since the number of rows of avengersWhoDied is 69, and the original dataframe, av, has 173 rows. **


FiveThirtyEight statement:
"Of the nine Avengers we see on screen — Iron Man, Hulk, Captain America, Thor, Hawkeye, Black Widow, Scarlet Witch, Quicksilver and The Vision — every single one of them has died at least once"

```{r}
#Luke Juffer
avengersWhoDied
heroes_to_exclude <- c("Anthony Edward \"Tony\" Stark", "Robert Bruce Banner", "Steven Rogers", "Thor Odinson", "Clinton Francis Barton", "Natalia Alianovna Romanova", "Wanda Maximoff", "Pietro Maximoff", "Victor Shade (alias)")

at_least_one_missing <- any(!(heroes_to_exclude %in% avengersWhoDied$Name.Alias))

if (at_least_one_missing) {
  print("At least one excluded hero is missing from avengersWhoDied.")
} else {
  print("All excluded heroes are present in avengersWhoDied.")
}


```

FALSE was returned, so all of the original nine avengers have died at least once.

Upload your changes to the repository. Discuss and refine answers as a team.


